---
title: "K-means CLustering Lab"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
library(readr)
library(png)
```

## Question 1 
**How many different types of hieroglyphics do you see?**
*Answered in the lab report.*

## Question 2
**Read in all the images in R and store them as a single data frame**
```{r}
files = list.files("./Hieroglyphics_Images")
hieroglyphs = lapply(files, function (x) readPNG(paste('./Hieroglyphics_Images/',x,sep = "")))
hieroglyphs_df <- data.frame(do.call(rbind, lapply(hieroglyphs, function(x) t(matrix(x)))))

write.csv(hieroglyphs_df, "hieroglyphs.csv") #save the image dataframe as a csv file
```

```{r}
hieroglyphs_df <- read.csv("hieroglyphs.csv") 
```

## Question 3
**Compress the data with PCA**
```{r}
hieroglyphs_df.pca <- prcomp(hieroglyphs_df, scale. = TRUE)

pca_var = hieroglyphs_df.pca$sdev^2          #calculate percentage of variance explained
prop_varex = pca_var/sum(pca_var)
variance_explained = cumsum(prop_varex)
plot(variance_explained, type = 'b')
variance_explained[2000]

hier_pca_df <- data.frame(hieroglyphs_df.pca$x)
write.csv(hier_pca_df, "hier_pca.csv")       #save the pca dataframe as a csv file
```
*Detailed answer provided in the lab report*
## Question 4
**Run some k-means clustering algorithm**
```{r}
clustering_model <- lapply(1:100, function (x) kmeans(hier_pca_df[,1:557],x,nstart=10, iter.max = 20))

wss <- lapply(clustering_model, function(x) sum(x$tot.withinss)) #wss of all 100 different values of k
bss <- lapply(clustering_model, function(x) sum(x$betweenss))   #bss of all 100 different values of k
wss_df = data.frame(k = 1:100)
bss_df = data.frame(k = 1:100)
wss_df$WSS <- wss
bss_df$BSS <- bss
wss_df$WSS <- as.numeric(wss_df$WSS)
bss_df$BSS <- as.numeric(bss_df$BSS)

#plot k from 1 to 100
plot(wss_df[1:100,'k'],wss_df[1:100,'WSS'], type = 'b', xlab = 'Number of Clusters', ylab = 'Within groups sum of squares')
plot(bss_df[1:100,'k'],bss_df[1:100,'BSS'], type = 'b', xlab = 'Number of Clusters', ylab = 'Between-cluster sum of squares')

#plot k from 1 to 60
plot(wss_df[1:60,'k'],wss_df[1:60,'WSS'], type = 'b', xlab = 'Number of Clusters', ylab = 'Within groups sum of squares')
plot(bss_df[1:60,'k'],bss_df[1:60,'BSS'], type = 'b', xlab = 'Number of Clusters', ylab = 'Between-cluster sum of squares')

write.csv(wss_df, "wss.csv")      #save the wss and bss data into csv file for future use if needed
write.csv(bss_df, "bss.csv")
```
*Detailed answer provided in the lab report*

## Question 5
**Perform the kmeans with your chosen k 1000 times, each time with 1 start**
```{r}
clustering_model2 <- lapply(1:1000, function (x) kmeans(hier_pca_df[,1:557],25,nstart=1, iter.max = 20))

wss <- lapply(clustering_model2, function(x) sum(x$tot.withinss))    #get the wss data of 1000 different random starts
wss_df_q5 = data.frame(k = 1:1000)
wss_df_q5$WSS <- wss
wss_df_q5$WSS <- as.numeric(wss_df_q5$WSS)

write.csv(wss_df_q5, "wss_q5.csv")        #save the wss data of 1000 nstarts as a csv file
ggplot(wss_df_q5, aes (WSS)) +
  geom_histogram() +
  labs(title = "Distribution of WSS given nstart = 1", x = "Within cluster sum of squares") +
  theme_bw() +
  theme(plot.title = element_text(hjust=0.5))
```
```{r}
clustering_model_final <- kmeans(hier_pca_df[,1:2000],25,nstart=20, iter.max = 20)
```
*Detailed answer provided in the lab report*

## Question 6
**Write each image to the folder corresponding to it's assigned cluster.**
```{r}
lapply(1:25, function (x) dir.create(paste("Cluster",x,sep = "_")))

image_num = 1
for (cluster_num in as.numeric(clustering_model_final$cluster)) {
  setwd(paste("Cluster",as.numeric(cluster_num),sep = "_"))
  writePNG(hieroglyphs[[image_num]], target = paste(image_num,".png",sep = ""))
  setwd("~/Library/Mobile Documents/com~apple~CloudDocs/Downloads/DA 350/clustering")
  image_num = image_num + 1
}
```
*Detailed answer provided in the lab report*

## Question 7
```{r}
#create a dataframe for the prediction set
files = list.files("./Predictions")
predictions = lapply(files, function(x) readPNG(paste('./Predictions/',x,sep = "")))
predict_df <- data.frame(do.call(rbind, lapply(predictions, function(x) t(matrix(x)))))

#scale the prediction set using loading vectors of traning data
predict_scaled <- data.frame(predict(hieroglyphs_df.pca, newdata = predict_df))
centers_df <- data.frame(clustering_model_final$centers)

predict_cluster <- function(new_data) {
  min_distance = dist(rbind(centers_df[1,],new_data))
  cluster = 1
  for (i in (2:25)) {
    distance = dist(rbind(centers_df[i,],new_data), method = "euclidean")
    if (distance < min_distance){
      min_distance = distance
      cluster = i
    }
  }
  return (cluster)
}
#predict
for (i in 1:10) {
  print(predict_cluster(predict_scaled[i,]))
}
```
*Detailed answer provided in the lab report*

